### 2020-06-29 paper.tex
initial commit

\documentclass{article}
\usepackage[utf8]{inputenc}
\usepackage[english]{babel}
\usepackage{alltt}
\usepackage{amsmath}
\usepackage{csquotes}
\usepackage{multicol}

\def\code#1{\texttt{#1}}

\usepackage{biblatex}
\addbibresource{sources.bib}

\evensidemargin=0in
\oddsidemargin=0in
\textwidth=6.5in
\textheight=9.0in

\title{Sorting by Reversals}
\author{Jonas Hübotter}

\begin{document}

\maketitle

\printbibliography

\end{document}