### 2020-07-03 paper.tex
update examples

\documentclass[11pt,DIV=11]{scrartcl}
\usepackage[utf8]{inputenc}
\usepackage[english]{babel}
\usepackage{alltt}
\usepackage{amsmath}
\usepackage{csquotes}
\usepackage{multicol}

\def\code#1{\texttt{#1}}

\usepackage[sorting=ynt,style=alphabetic]{biblatex}
\addbibresource{sources.bib}

\title{Sorting by Reversals}
\author{Jonas Hübotter}

\begin{document}

\maketitle

\printbibliography

\end{document}